Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2019-02-25 draft.tex
small fixes

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ of the preference of an agent $j$. 
The adversary has to make the score of $y$ as high as possible, and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \ppref_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. 
Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \ppref_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \ppref_j y$). 
In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference in scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of score is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:
\begin{align} 
\label{eq:compl}
\begin{split}
a \pref_j x &⇔ ¬(x \ppref_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \ppref_j y) ∨ (x \ppref_j a)).
\end{split}
\end{align} 
In what follows let $\hat{\profile}$ be the profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies (\ref{eq:compl}).
%There exists a completion ${\pref_j} \in C(\ppref_j)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile), {\pref_j} \in \profile} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that satisfies 
\end{claim}
%In this claim, we abuse notation by writing ${\pref_j} \in \profile$ to mean that the preference ordering corresponding to agent $j$ in $\profile$ is $\pref_j$.
\noindent It follows that $\hat{v}_{j}(x)$, the rank of $x$ for voter $j$ in the PMR-maximizing linear order $\bar{v}_{j}$ %$\PMR(x,y; \pprofile, W)$ 
is $\hat{v}_{j}(x) = 1+\card{A}-\card{{\ppref_j}(x)}$, where ${\ppref_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\ppref_j$, and the rank of $y$ is $\hat{v}_{j}(x)=1+\card{{\pprefinv_j}(y)}+\beta$, where $\beta = \card{A} - \card{{\ppref_j}(x)}$ if $(x \ppref_j y)$ and $\alpha = 0$ otherwise.

@@

Observe that the question may be defined equivalently as whether $q w_{r} + p w_{r+2} ≥ (p + q) w_{r+1}$, where $p, q$ are natural numbers. 
As a first attempt to make this question concrete, let us define a profile $P$ containing $p+q$ agents. 
Define an alternative $a$ that receives $q$ times the rank $r$ and $p$ times the rank $r+2$ and an alternative $b$ that receives $p+q$ times the rank $r+1$. 
Observing that the score of $a$ in that profile is exactly the left hand side of the question, and that the score of $b$ is the right hand side, intuition suggests that observing whether the chair picks $a$ or $b$ as winner will let us determine which side is greater; equality occurring when the chair declares $a$ and $b$ as tied for the victory. 
However, we still need to complete the profile, thus, come up with other $m-2$ alternatives defined so that each agent in the resulting profile has placed exactly one alternative in each rank. 
And we must ensure, doing this, that these other alternatives are not better than $a$ or $b$.\footnote{Because, if the chair picks a different alternative $c$, it only tells us that the chair prefers $c$ to both $a$ and $b$, but it generally does not answer the question we are interested in.} 
Taking $p = q = 1$, $m=4$, $r=2$, we see that such a completion may be impossible. Luckily, this problem can be worked around, at the price of increasing the number of agents.

>>>>

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ of the preference of an agent $j$. 
The adversary has to make the score of $y$ as high as possible, and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \ppref_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. 
Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \ppref_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \ppref_j y$). 
In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference in scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of score is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:
\begin{align} 
\label{eq:compl}
\begin{split}
a \pref_j x &⇔ ¬(x \ppref_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \ppref_j y) ∨ (x \ppref_j a)).
\end{split}
\end{align} 
In what follows let $\hat{\profile}$ be the profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies (\ref{eq:compl}).
%There exists a completion ${\pref_j} \in C(\ppref_j)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} \max_{\profile \in C(\pprofile), {\pref_j} \in \profile} [ s(y; \profile,\w) - s(x; \profile,\w) ]$ and that satisfies 
\end{claim}
%In this claim, we abuse notation by writing ${\pref_j} \in \profile$ to mean that the preference ordering corresponding to agent $j$ in $\profile$ is $\pref_j$.
\noindent It follows that $\hat{v}_{j}(x)$, the rank of $x$ for voter $j$ in the PMR-maximizing linear order $\bar{v}_{j}$ %$\PMR(x,y; \pprofile, W)$ 
is $\hat{v}_{j}(y) = 1+\card{A}-\card{{\ppref_j}(x)}$, where ${\ppref_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\ppref_j$, and the rank of $y$ is $\hat{v}_{j}(x)=1+\card{{\pprefinv_j}(y)}+\beta$, where $\beta = \card{A} - \card{{\ppref_j}(x)}$ if $(x \ppref_j y)$ and $\beta = 0$ otherwise.

@@

Observe that the question may be defined equivalently as whether $q w_{r} + p w_{r+2} ≥ (p + q) w_{r+1}$, where $p, q$ are natural numbers. 
As a first attempt to make this question concrete, let us define a profile $P$ containing $p+q$ agents. 
Define an alternative $a$ that receives $q$ times the rank $r$ and $p$ times the rank $r+2$ and an alternative $b$ that receives $p+q$ times the rank $r+1$. 
Observing that the score of $a$ in that profile is exactly the left hand side of the question, and that the score of $b$ is the right hand side, intuition suggests that observing whether the chair picks $a$ or $b$ as winner will let us determine which side is greater; equality occurring when the chair declares $a$ and $b$ as tied for the victory. 
However, we still need to complete the profile, thus, come up with other $m-2$ alternatives defined so that each agent in the resulting profile has placed exactly one alternative in each rank. 
And we must ensure, doing this, that these other alternatives are not better than $a$ or $b$.\footnote{Because, if the chair picks a different alternative $c$, it tells us that the chair prefers $c$ to both $a$ and $b$, 
%but it generally does not  answer the question we are interested in
but in general we do not acquire the type of constraint we are interested in.} 
Taking $p = q = 1$, $m=4$, $r=2$, we see that such a completion may be impossible. Luckily, this problem can be worked around, at the price of increasing the number of agents.
